---
title: "Motor Trend Car Road Tests"
author: "Fernanda Sánchez"
date: "10/12/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Reproducible Pitch Presentation

### Cars Consumption: comparation using Milles per Galon (MPG)

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

### Source
Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391-411.
```{r}
library(datasets)
library(shiny)

head(mtcars, 3)
```

## Format

A data frame with 32 observations on 11 (numeric) variables.

[, 1]	mpg	Miles/(US) gallon

[, 2]	cyl	Number of cylinders

[, 3]	disp	Displacement (cu.in.)

[, 4]	hp	Gross horsepower

[, 5]	drat	Rear axle ratio

[, 6]	wt	Weight (1000 lbs)

[, 7]	qsec	1/4 mile time

[, 8]	vs	Engine (0 = V-shaped, 1 = straight)

[, 9]	am	Transmission (0 = automatic, 1 = manual)

[,10]	gear	Number of forward gears

[,11]	carb	Number of carburetors


## Analysis - Main Code

  
    paste("mpg ~", "as.integer(", input$variable, ")")
  
  
    lm(as.formula(formulaTextPoint()), data=mpgData)  

  
    summary(fit()) })
  
  output$mpgPlot <- renderPlot({
  
    with(mpgData, {
    
      plot(as.formula(formulaTextPoint()))
      
      abline(fit(), col=2)
      
    })  })


